---
title: "Seabirds"
output:
  html_document:
    toc: true
    toc_float: true
    number_sections: true
    df_print: paged
    css: ../../../styles.css
  pdf_document: default
---
```{r include = FALSE}
library(tidyverse)
library(here)

seabirds <- read_csv(here::here("clean_data/seabirds.csv")) # load data
```
# About the task
The data provided has details of seabird sightings recorded from mid 1969
to the end of 1990. Information of the birds sighted is held on a separate sheet
from the information about the location, date and time of the sightings.  To
investigate the data only a small selection of columns are needed from the
detailed records.  

The "count" data (the number of birds seen in an observation) does not seem
particularly reliable, with lots of NA values and some large estimates that go
as high as 99999. For that reason, some of the conclusions involving total 
numbers of birds should be treated with caution.
<br>
<br>
  
# Packages and Data Clean
The following packages were loaded: tidyverse, here, readxl  

For the analysis, from the bird data, only details about the record ID, bird
species and count were needed. From ship data, only the record ID and latitude
of the ships were needed. The record ID is a primary key in the ship data and a
foreign key in the bird data, leading to a one-to-many relationship when the
data is joined.
<br>
<br>

# Assumptions
In the count column there are around 200 observations that exceed 500 birds,
with some counts as high as 99999. I chose to make these counts NA, as I didn't
want wild estimates throwing off my analysis. The presence of these large
numbers makes me doubt the validity of some of the data. I would like more
context as to how these large estimates are arrived at.  

There are NA values in the count column. Some of these appear to relate to days
when no observations of birds occurred. However, others are observations of a
particular species where, presumably, the observer has forgotten to enter a
number of birds or been unable to count them. These observations have been left
in the data as they are an observation of a particular species on a particular
day, but the NA values will need to be skipped when performing calculations.

The latitude column also contains NA values. This will affect he result of
question 3.
<br>
<br>

# Analysis
## Most Sightings
Which bird had the most individual sightings?
```{r message=FALSE}
seabirds %>% 
  group_by(common_name, sci_name, abbreviation) %>%
  summarise(sightings = n()) %>% 
  ungroup() %>% 
  slice_max(sightings, n = 3)
```
The most individual sightings were of the wandering albatross sensu lato PL2, 
scientific name *Diomedea antipodensis / exulans PL2*.
<br>

## Most Birds
Which bird had the highest total count?
```{r message=FALSE}
seabirds %>% 
  group_by(common_name, sci_name, abbreviation) %>%
  summarise(total_count = sum(count, na.rm = TRUE)) %>% 
  ungroup() %>% 
  slice_max(total_count, n = 3)
```
After excluding any count that exceeded 500, the highest count for any single
species was the prion, scientific name *Pachyptila*.  It should be noted that
this total likely still contains estimated numbers so may not be entirely
reliable.
<br>

## Most Birds Above -30 Latitude
Which bird had the highest total count above a latitude of -30?
```{r message=FALSE}
seabirds %>% 
  filter(lat > -30) %>% 
  group_by(common_name, sci_name, abbreviation) %>%
  summarise(total_count = sum(count, na.rm = TRUE)) %>% 
  ungroup() %>% 
  slice_max(total_count, n = 3)
```
Above a latitude of -30 degrees, the wedge-tailed shearwater,
*Puffinus pacificus*, had the highest total count.
<br>

## Solo Birds
How many different types of birds were only ever seen in groups of one?
```{r}
seabirds %>% 
  group_by(common_name) %>%
  filter(sum(count, na.rm = TRUE) == n()) %>% 
  distinct(common_name) %>% 
  nrow()
```
78 distinct species of bird were only observed alone.
<br>

## Penguins
How many penguins were seen?
```{r}
seabirds %>%
  filter(str_detect(common_name, "(?i)penguin")) %>% 
  summarise(total_count = sum(count, na.rm = TRUE))
```
In total, 158 penguins were seen.