---
title: "Cake Ingredients"
output:
  html_document:
    toc: true
    toc_float: true
    number_sections: true
    df_print: paged
    css: ../../styles.css
  pdf_document: default
---
```{r include = FALSE}
library(tidyverse)
library(here)

cake_ingredients <-
  read_csv(here::here("clean_data/cake_ingredients_clean.csv")) # load data
```
# About the task
The data sets have ingredient lists for a number of cake recipes. The data is
split between two csv files, one containing types of cake, codes for recipe
ingredients and number of units, the other containing the ingredient codes,
their full name and the measurement units. The data sets need joined in an
appropriate way before analysis can take place.
<br>
<br>

# Packages and Data Clean
The following packages were used: tidyverse, here, janitor  
  
The cake ingredients were pivoted longer to make an easier-to-use list.
Ingredients that were NA in a recipe were dropped from the list. The two data 
sets were joined using the ingredient code as a primary key. The resulting data
frame was tidied, checked and saved into the clean data folder.
<br>
<br>

# Assumptions
Once tidied and joined, the data is complete.  

For the purposes of analysis, egg whites, egg yolks and eggs are considered
separate ingredients. Motivation for this is that if a recipe asked for egg
whites then you are unlikely to get the same result using whole eggs.  

Some measurements provided for cup sizes have been rounded up or down.  This
does not affect the analysis, but may be critical for the recipe. For example,
0.3 cups does not exist as a standard measure, but is it 1/4 cup rounded up or
1/3 cup rounded down?
<br>
<br>

# Analysis
## Cocoa Content
Which cake has the most cocoa in it?
```{r}
cake_ingredients %>% 
  filter(ingredient == "Cocoa") %>% 
  slice_max(amount, n = 3)
```
Only two recipes feature cocoa. One Bowl Chocolate contains the most.
<br>

## Sponge Cake
For sponge cake, how many cups of ingredients are used in total?
```{r}
cake_ingredients %>% 
  filter(cake == "Sponge")

cake_ingredients %>% 
  filter(cake == "Sponge", measure == "cup") %>% 
  summarise(sum(amount))
```
The are 3.5 *cups* of ingredients in the sponge cake recipe. There are also four
eggs and some ingredients that are measured by teaspoons.
<br>

## Teaspoons
How many ingredients are measured in teaspoons?
```{r}
cake_ingredients %>% 
  filter(measure == "teaspoon") %>% 
  distinct(ingredient) %>% 
  summarise(measured_by_teaspoon = nrow(.))

cake_ingredients %>% 
  filter(measure == "teaspoon") %>% 
  summarise(measured_by_teaspoon = nrow(.))
```
Eight distinct ingredients are measured by teaspoon. These ingredients feature
in multiple recipes so that there are 45 entries in the whole table measured by
teaspoons.
<br>

## Unique Ingredients
Which cake has the most unique ingredients?  

There are two ways to interpret this query. Firstly, which cake contains the
most different (unique) ingredients.
```{r}
cake_ingredients %>% 
  count(cake) %>% 
  slice_max(n, n = 3)
```
Babas au Rhum and One Bowl Chocolate each have 11 distinct ingredients.

The second way to interpret the query is that it is asking which recipes
contain ingredients that only appear in that recipe (unique), and then which
recipe contains the most of these unique ingredients.
```{r}
rare_ingredients <- cake_ingredients %>%
  count(ingredient) %>% 
  filter(n == 1) 

inner_join(rare_ingredients, cake_ingredients, by = "ingredient") %>% 
  select(ingredient, cake) %>% 
  arrange(cake)
```
Angel cake is the only recipe that contains two ingredients that are unique
to that recipe. These ingredients are cream of tartar and egg white.
<br>

## Rare Ingredients
Which ingredients are used only once?
```{r}
cake_ingredients %>%
  count(ingredient) %>% 
  filter(n == 1)
```
Eight ingredients are only used once in this selection of recipes:  
* Bananas  
* Cream of tartar  
* Crushed Ice  
* Dried currants  
* Egg white  	
* Nutmeg  	
* Nuts  	
* Zwiebach