---
title: "Hallowe'en Candy"
output:
  html_document:
    toc: true
    toc_float: true
    number_sections: true
    df_print: paged
    css: ../../../styles.css
  pdf_document: default
---
# About the Task
Since 2006 David Ng and Ben Cohen have conducted a light-hearted ranking of
Hallowe'en candy. At first they focused on personal, anecdotal experience. What 
started as an amusing exercise for their blog ballooned into a quasi-scientific
analysis of public candy perceptions. In this task, I have taken three years of
their data and will analyse some sweeping generalisations that can be made from
this not-so-reliable cross-section of society.
<br>
<br>

# Packages
The following packages were used: read_xl, tidyverse, here, janitor. The 
package plyr was used for one question only.
<br>
<br>

# Data Clean  
The initial clean removed unwanted columns from the data sets. The candy data
contained a number of questions that were comic references or not candy at all;
these columns were excluded (see *Assumptions* for more detail). Column titles
were tidied up and an id assigned to each observation. Three years of
observations were combined and a further sweep of the columns was carried out to
ensure that matching candies had been combined where possible.  
  
Character strings were removed from respondents' ages. Some rows had ages that
were outside a reasonable range, these were set to NA (see *Assumptions* for
more detail). The column was transformed to a numerical type.
  
In 2016 and 2017, respondents were asked to give their country. Where people had
responded with fictitious countries or joke answers, their responses were set to
NA. Wherever possible, a real country or area was left in the data  (see 
*Assumptions* for more detail). All variations of USA, Canada and UK were
collected together.  

A short script at the end assigns point values to the candy ratings for the
questions in section 6
<br>
<br>

# Assumptions  
## Candy
The following columns were excluded from the analysis as they are not candy. The
reason, if needed, *is in italics*.  
* Abstained from M&M'ing, *joke answer following lots of M&M types*  
* Bonkers (the board game)  
* Box'o'Raisins, *fruit*  
* Broken glow stick  
* Cash, or other forms of legal tender  
* Chardonnay  
* Creepy Religious comics/Chick Tracts, *comic*  
* Dental paraphenalia  
* Generic Brand Acetaminophen, *medicine*  
* Glow sticks  
* Healthy Fruit  
* Hugs (actual physical hugs)  
* JoyJoy (Mit Iodine), *ficticious, from the Simpsons*  
* Kale smoothie  
* Lapel Pins  
* Minibags of chips, *crisps*  
* Mint Leaves, *no evidence of mass-marketed candy of this name*  
* Peanut Butter Jars  
* Pencils  
* Person of Interest Season 3 DVD Box Set  
* Peterson Brand Sidewalk Chalk  
* Real Housewives of Orange County Season 9 Blue-Ray  
* Sandwich-sized bags filled with BooBerry Crunch, *cereal*  
* Senior Mints, *senior mints exist only as a joke product*  
* Spotted Dick, *dessert*  
* Sweetums, *ficticious, from Parks & Rec*  
* Vials of pure high fructose corn syrup, for main-lining into your vein  
* Vicodin, *medicine*  
* White Bread  
* Whole Wheat anything  
Open ended questions allowing further examples of candy that elicits "joy" or
"despair" were discounted.  
The survey concluded with a number of topical or joke questions that were also
excluded for not being candy related.  

Despite concerns, the following candies were allowed to remain. Due to the age
of some respondents, discontinued candy was allowed to remain.  
* Reggie Jackson Bar, *discontinued 1981*  
* Trail Mix, *usually contains candy*  
* Any full-sized candy bar, *generic but valid*  
* Anonymous brown globs that come in black and orange wrappers, *some respondents knew what these were*  
* Bonkers, *discontinued 1990s*  
* Goo Goo Clusters, *niche but definitely candy*
<br>

## Ages
When tidying respondents' ages, the following assumptions were made:  
* character strings and punctuation were removed (resulting blanks become NA)  
* under 5 set to NA (assume no respondent was genuinely under 5 years old)  
* over 117 set to NA (the oldest person alive at the time was 117)  
This left 441 respondents with no age. It was noted that there were still 18
respondents over the age of 80, with one claiming to be 115 years old.
<br>

## Countries
After tidying answers, the following assumptions were applied.  

### Numerical, Ficticious Countries or Joke Answers
Numerical answers were set to NA during the initial tidy. The following
responses were deemed equivalent to not answering the question and changed to NA.  
a, atlantis, denial, earth, fear_and_loathing, god (came from "god's country"),
i_don (came from "i don't know anymore"), insanity_lately, narnia, neverland,
one_of_the_best_ones, see_above, somewhere, subscribe_to_dm (came from 
"subscribe to dm4uz3 on youtube), there_isn (came from "there isn't one for old
men"), "this_one".  
  
### Canada, USA, UK and other
The main question that involved countries was about Canada, USA and UK. At that
time, all other non-NA answers would be grouped as "other".  
  
The following answers were deemed to mean Canada:  
can, i_pretend_to_be_from_canada, soviet_canuckistan, canada.  
  
The following answers were grouped as UK:  
endland, england, scotland, united_kindom, united_kingdom, uk.  
  
The following mess was set to be USA:  
ahem (came from "ahem....amerca"), alaska, america, california, eua, merica,
murica, murrika, new_jersey, new_york, north_carolina, pittsburgh, sub (came
from "sub-canadian north america... 'merica"), the_best_one_ (came from "the
best one - usa"), the_yoo_ess_of_aaayyyyyy, trumpistan, u_s, u_s_a, ud,
unhinged_states, unied_states, unite_states, united_sates, united_staes, 
united_state, united_statea, united_stated, united_statss, united_stetes,
united_ststes, unites_states, units_states, us_of_a, usa_usa_usa,
usa_usa_usa_usa, usaa, usas, usausausa, ussa, the_united_states,
the_united_states_of_america, united_states, united_states_of_america, us, usa.  
  
Any other answer that was deemed to be a real geographical area was left to be
grouped into the "other" category.  These included such nuggets as:  
a_tropical_island_south_of_the_equator, canae (in Turkey), europe, n (came from
n.america), not_the_us_or_canada.  
A few respondents answered cascadia or the_republic_of_cascadia, which is a
region in western North America extending across the US/Canada border. It was
deemed that these respondents would not want to be grouped with either US or
Canada so were left separate.  
<br>
<br>

```{r include = FALSE}
library(tidyverse)
library(here)

all_candy <- read_csv(here::here("clean_data/all_candy.csv")) # Clean data
```
# Analysis
## Total Candy Ratings
What is the total number of candy ratings given across the three years.
```{r}
all_candy %>% 
  select(any_full_sized_candy_bar:york_peppermint_patties) %>% # columns containing candy data
  summarise(num_ratings = sum(!is.na(.)))
```
Once non-candies have been removed, there are 583908 candy ratings from the
three years of data.
<br>

##  Average Age of Trick-or-Treaters
What was the average age of people who are going out trick or treating?
```{r}
all_candy %>% 
  select(age, going_out) %>% 
  group_by(going_out) %>% 
  summarise(mean_age = round(mean(age, na.rm = TRUE), 1),
            median_age = median(age, na.rm = TRUE))
```
The mean age of survey respondents who said they were going trick-or-treating
was 35.2 years old. The median age was 37.
<br>

## Average Age of Non-Trick-or-Treaters
What was the average age of people who are not going trick or treating?  
  
The mean age of survey respondents who said they were not going
trick-or-treating was 39.2 years old. The median age was 37.5.
<br>

## Ratings Summary
For each of joy, despair and meh, which candy bar received the most of these
ratings?  

```{r}
all_candy %>% 
  select(any_full_sized_candy_bar:york_peppermint_patties) %>% 
  plyr:: ldply(function(x) sum(x == "JOY", na.rm = TRUE)) %>% 
  rename(candy = .id, amount_of_joy = V1) %>% 
  arrange(desc(amount_of_joy)) %>% 
  head(3)

all_candy %>% 
  select(any_full_sized_candy_bar:york_peppermint_patties) %>% 
  plyr::ldply(function(x) sum(x == "MEH", na.rm = TRUE)) %>% 
  rename(candy = .id, amount_of_meh = V1) %>% 
  arrange(desc(amount_of_meh)) %>% 
  head(3)

all_candy %>% 
  select(any_full_sized_candy_bar:york_peppermint_patties) %>% 
  plyr::ldply(function(x) sum(x == "DESPAIR", na.rm = TRUE)) %>% 
  rename(candy = .id, amount_of_despair = V1) %>% 
  arrange(desc(amount_of_despair)) %>% 
  head(3)
```
Full sized candy bars elicited the most "joy" from participants of the survey.
The most joy from a specific candy was Reece's Peanut Butter Cups.  
More participants thought lollipops were "meh" than any other candy.  
Gum from baseball cards brought "despair" from the most people.
<br>

## Starburst Despair
How many people rated Starburst as despair?
```{r}
count(all_candy, starburst)
```
1990 people rated starburst as despair.
<br>
<br>

# Analysis using points allocation
For the remaining queries, candy ratings have been assigned as points:  
* JOY is worth +1  
* MEH is worth 0  
* DESPAIR is worth -1
```{r include = FALSE} 
all_candy_points <- read_csv(here::here("clean_data/all_candy_points.csv"))
```
<br>

## Most Popular Candy by Gender
What was the most popular candy bar for each gender in the dataset?
```{r}
all_candy_points %>% 
  group_by(gender) %>% 
  summarise(across(any_full_sized_candy_bar:york_peppermint_patties, sum)) %>% 
  pivot_longer(cols = any_full_sized_candy_bar:york_peppermint_patties,
               names_to = "candy") %>% 
  group_by(gender) %>% 
  slice_max(value, n = 2)
```
In every gender category full sized candy bars were the most popular hallowe'en
candy. The most popular specific candies were Reese's Peanut Butter Cups for
respondents who identified themselves as female, male and those who didn't
answer the question. Those who selected "I'd rather not say" liked KitKats; 
"Other" enjoyed Twixes.  
Note that gender was not asked for in the 2015 survey, resulting in a large
number of responses in NA.
<br>

## Most Popular Candy by Year
What was the most popular candy bar in each year?
```{r}
all_candy_points %>% 
  mutate(year_id = str_sub(year_id, 1, 4)) %>% 
  group_by(year_id) %>% 
  summarise(across(any_full_sized_candy_bar:york_peppermint_patties, sum)) %>% 
  pivot_longer(cols = any_full_sized_candy_bar:york_peppermint_patties,
               names_to = "candy") %>% 
  group_by(year_id) %>% 
  slice_max(value, n = 2)
```
Each year, full sized candy bars proved to be most popular with respondents. In
terms of a specific type of of candy, Reese's Peanut Butter Cups took the crown
in 2015 and 2017. KitKat was most popular in 2016.

## Most Popular Candy by Country
What was the most popular candy bar for people in US, Canada, UK, and all other countries?
```{r}
all_candy_points %>% 
  drop_na(country) %>% 
  mutate(country = case_when(country == "USA" ~ "USA",
                             country == "CAN" ~ "Canada",
                             country == "GBR" ~ "UK",
                             .default = "other")
         ) %>% 
  group_by(country) %>% 
  summarise(across(any_full_sized_candy_bar:york_peppermint_patties, sum)) %>% 
  pivot_longer(cols = any_full_sized_candy_bar:york_peppermint_patties,
               names_to = "candy") %>% 
  group_by(country) %>% 
  slice_max(value, n = 2)
```
In each of the country categories, full sized candy bars were most popular.  
Specific types of candy that were most popular were:  
* Reese's Peanut Butter Cups in the USA  
* Lindt Truffles, Rolos and Toblerone (three-way-tie) for the UK  
* KitKat in Canada and in "other countries"  
Note that country was not asked for in the 2015 survey, resulting in a large
number of NA being dropped from the reults.