---
title: "Decathlon"
output:
  html_document:
    toc: true
    toc_float: true
    number_sections: true
    df_print: paged
    css: ../../../styles.css
  pdf_document: default
---
# About the task
A data set was provided that contained results from two decathlon competitions, 
Decastar and the Olympic Games. Times and distances for each of the ten events
were given, as well as each competitor's final points total and rank in that
competition. Not all athletes competed in both competitions so any totaling of 
points across the competitions may be skewed by that.
<br>
<br>

# Packages and Data Clean
The following packages were used: tidyverse, here, janitor  
  
The initial clean tidied text strings and column titles and created a separate
column for the competitor's names. The data contained no missing information and
no assumptions needed to be made.  
  
There may be some merit in pivoting the table to be longer, but for the analysis
questions that are given this will not be necessary.
<br>
<br>

```{r include = FALSE}
library(tidyverse)
library(here)

decathlon_data <- read_csv(here::here("clean_data/decathlon_clean_data.csv"))
```
# Analysis
## Best Long Jump
Who had the longest long jump seen in the data?
```{r}
decathlon_data %>% 
  select(name, long_jump) %>% 
  slice_max(long_jump, n = 3)
```
The competitor named Clay had the longest long jump, at 7.96 metres.
<br>

## Average 100 metre Times
What was the average 100m time in each competition?
```{r}
decathlon_data %>% 
  select(competition, x100m) %>% 
  group_by(competition) %>% 
  summarise(average_100m_time = round(mean(x100m), 3))
```
The average time for the 100m at Decaster was 11.175 seconds. At the Olympic
games the average time was faster, at 10.916 seconds.
<br>

## Highest Points Total
Who had the highest total points across both competitions?
```{r}
# Reading question as the totals from the two competitions combined
decathlon_data %>% 
  select(name, points, competition) %>% 
  group_by(name) %>% 
  summarise(total_points = sum(points)) %>% 
  slice_max(total_points, n = 3)
```
Sebrle had the highest combined total from the two competitions. It should be
noted than not all competitors competed at both competitions.
<br>

## Best Competitors' Shot-Put
What was the shot-put scores for the top three competitors in each competition?
```{r}
decathlon_data %>% 
  select(name, shot_put, points, competition) %>% 
  group_by(competition) %>% 
  slice_max(points, n = 3) # Top three competitors by points total
```
The top three points scorers at each competion are Sebrle, Clay and Karpov, in
that order.  
Their shot-put distances at Decaster were: 14.83, 14.26 and 14.77 metres
respectively.  
Their shot put distances at the Olympic Games were: 16.36, 15.23 and 15.93
metres respectively.
<br>

## Fast 400m
What was the average points for competitors who ran the 400m in less than 50
seconds vs. those than ran 400m in more than 50 seconds?
```{r}
decathlon_data %>% 
  select(x400m, points) %>% 
  mutate(fast_400m = case_when(x400m <= 50 ~ TRUE,
                               x400m > 50 ~ FALSE)) %>% 
  group_by(fast_400m) %>% 
  summarise(mean_points = round(mean(points), 0))
```
On average, competitors who ran the 400 metres in 50 seconds or under scored
almost 400 points more across the whole competition than those who did not.